---
layout: 12pt
output: pdf_document
bibliography: sagecast.bib
csl: ecology.csl
header-includes:
 \usepackage{float}
 \floatplacement{figure}{H}
---

\begin{centering}
\textbf{\Large{Supporting Information for:\\Dynamic spatiotemporal modeling of a habitat defining plant species to support wildlife management at regional scales}}

\textsc{\small{Andrew T. Tredennick, Thomas Prebyl, Adrian P. Monroe, John Lombardi, and Cameron L. Aldridge}}
\end{centering}


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE)
library(readxl)
library(dplyr)
library(ggplot2)
library(tidyverse)
```


# Potential scale reduction factors

Potential scale reduction factors ($\hat{R}$) help diagnose MCMC convergence.
MCMC algorithms have reached convergence when $\hat{R} < 1.1$.

```{r rhats, eval = TRUE, results = "asis"}
rhats <- readRDS("../Output/rhats.RDS")
cores <- names(rhats)
for(doCore in cores) {
   tab <- rhats[[doCore]]$psrf
   print(knitr::kable(tab, digits = 2, format = "latex", caption = doCore))
}
```

\pagebreak

# Bayesian p-values
```{r, eval = TRUE, results = "asis"}
pvals <- readRDS("../Output/BayesianPValues.RDS") %>%
  drop_na()
news <- readRDS("../Output/BayesianPValues-Redos.RDS") %>%
  drop_na()
allp <- bind_rows(pvals, news)
colnames(allp) <- c("P-value type", "P-value", "Core Area")
knitr::kable(allp, digits = 2)
```

\pagebreak

# Posterior distributions

```{r posteriors, eval = TRUE, results = 'asis'}
mcmcDir <- "../Output/MCMC/"
subDirs <- list.files(mcmcDir)
mcmcList <- list()

paramLabs <- c("Intercept", "Dens. dep.", "Precip.", "Temp.", "Temporal std. dev.")
paramLabs <- factor(paramLabs, levels = paramLabs)

tableLabs <- c("Intercept, $\\beta_{1}$", 
               "Density dependence, $\\beta_2$", 
               "Precipitation effect, $\\beta_3$", 
               "Temperature effect, $\\beta_4$", 
               "Std. dev. of temporal random effect, $\\sigma_y$")
# tableLabs <- factor(tableLabs, levels = tableLabs)
    
for(doArea in subDirs) {
   tmp <- readRDS(paste0(mcmcDir, doArea, "/MCMC.RDS"))
   mcmcList[[doArea]] <- tmp
   tmpDf <- tmp[, c("Beta[1]", "Beta[2]", "Beta[3]", "Beta[4]", "sigma_y")]
   colnames(tmpDf) <- paramLabs[1:5]
   tmpDf <- as_tibble(tmpDf) %>%
      mutate(iteration = 1:n()) %>%
      gather(key = "Parameter", value = "value", -iteration) %>%
      mutate(Parameter = factor(Parameter, levels = paramLabs))
   
   p <- ggplot(tmpDf, aes(x = value, y=..density..)) +
    geom_histogram(color = "black", fill = "grey") +
    facet_wrap(~Parameter, nrow = 2, scales = "free") +
    ggthemes::theme_few(base_size = 12) +
    labs(x = "Parameter estimate", y = "Posterior density") +
    ggtitle(doArea)
   
   print(p)
   
   # table of posterior summaries
   postSummary <- tmpDf %>%
     group_by(Parameter) %>%
     summarise(Mean = mean(value),
               Median = median(value),
               SD = sd(value),
               `2.5%` = quantile(value, 0.025),
               `97.5%` = quantile(value, 0.975)) %>%
     dplyr::select(-Parameter) %>%
     as.data.frame()
   rownames(postSummary) <- tableLabs
     
   
   print(knitr::kable(postSummary, digits = 2))
   cat("\n")
}
```

# Estimated equilibrium cover

```{r equilibrium-cover, fig.cap="Posterior distributions of equilibrium cover calculated from fitted model parameters. The vertical black lines show the observed mean cover for each core area from 1985-2018."}
obscover <- read.csv("../Output/historic_Sage_Cover_Summary.csv") %>%
  group_by(coreName) %>%
  summarise(MeanCover = mean(mean)) %>%
  mutate(NameNoSep = gsub(" ", "", coreName)) %>%
  left_join(nameMap) %>%
  filter(!Name %in% badSites)

allpostdirs <- list.files("../Output/MCMC/")
allposts <- tibble()
for(d in allpostdirs) {
  f <- paste0("../Output/MCMC/", d, "/MCMC.RDS")
  if(file.exists(f)) {
    tmp <- readRDS(f)
    cover <- exp(tmp[ , "Beta[1]"] / (1 - tmp[,"Beta[2]"]))
    df <- data.frame(CoreArea = d,
                     Cover = cover)
    allposts <- bind_rows(allposts, df)
  }
}

allposts <- allposts %>%
  rename("NameNoSep" = CoreArea) %>%
  left_join(nameMap) %>%
  filter(!Name %in% badSites)

ggplot(allposts, aes(x = Cover)) +
  geom_density(color = NA, fill = "tan", adjust = 2) +
  geom_vline(data = obscover, aes(xintercept = MeanCover)) +
  facet_wrap(~Abbreviation, scales = "free") +
  labs(x = "Average equilibrium cover (%)", y = NULL) +
  theme(axis.line.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  scale_x_continuous(breaks = scales::pretty_breaks(3))
```


# Estimated colonization probabilities

```{r colonization-res}
colonization <- readRDS("../Output/ColonizationModelResults.RDS") %>%
  rename("Core area" = CoreArea,
         "Pr(colonize | cover = 0)" = ProbabilityOfColonization,
         "Mean cover in colonized cells" = MeanColonizationCover) %>%
  filter(!`Core area` %in% badSites)

kable(colonization, 
      caption = "Results from colonization model for each core area. Pr(colonize | cover = 0) reads, 'the probability of colonizition given that current cover is zero.'",
      digits = 2,
      label = "colonization",
      format = "latex", 
      booktabs = TRUE,
      linesep = "",
      escape = FALSE) 
```

<!-- # Temporal random effects -->

```{r temporals, eval = FALSE}
allGammas <- tibble()
for(doSite in subDirs) {
  tmp <- mcmcList[[doSite]]
  tmp <- as.data.frame(tmp) %>%
     mutate(iteration = 1:n()) %>%
     gather("parameter", "value", -iteration) %>%
     filter(grepl("gamma", parameter)) %>%
     group_by(parameter) %>%
     summarise(Mean = mean(value),
               Lower95 = quantile(value, 0.025),
               Upper95 = quantile(value, 0.975),
               .groups = "drop") %>%
     mutate(Site = doSite)
  allGammas <- bind_rows(allGammas, tmp)
}

allGammas <- allGammas %>%
   separate(parameter, into = c("out", "tmp"), sep = "\\[") %>%
   separate(tmp, into = c("yearid", "out2"), sep = "\\]") %>%
   dplyr::select(-out, -out2) %>%
   mutate(yearid = as.numeric(yearid) + 1985) %>%
   mutate(yearid = ifelse(yearid > 2012, yearid + 1, yearid)) %>%
   mutate(Site = ifelse(Site == "NLar", "North Laramie", Site))

ggplot(allGammas, aes(x = yearid)) +
   geom_ribbon(aes(ymin = Lower95, ymax = Upper95), alpha = 0.2) +
   geom_line(aes(y = Mean)) +
   geom_point(aes(y = Mean)) +
   geom_hline(aes(yintercept = 0), linetype = 2) +
   ggthemes::theme_few() +
   labs(x = "Year", y = "Intercept offset") +
   facet_wrap(~Site, scales = "free_y") +
   scale_x_continuous(breaks = seq(min(allGammas$yearid), max(allGammas$yearid), by = 2)) +
   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```


# Nesting and summer cover thresholds

```{r targs,  eval = TRUE, results = 'asis'}
read.csv("../AuxData/CoreAreaNamesMapping.csv") %>%
  dplyr::select(Name, Abbreviation, Region, NestingTarget, SummerTarget) %>%
  knitr::kable() %>%
  print()

```
